---
title: "Published data sets"
output: html_notebook
---

This document details the analysis of data from published studies using the Easy nanoLC and the Evosep One. 

<br>

### Data analysis and plotting

<br>

These are the R libraries we will require.

```{r, results = 'hide'}
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(viridis)
library(pheatmap)
library(dplyr)
library(data.table)
library(reshape2)
```


<br>

First we can plot the chromatograms. 

```{r}
###################################################################################################################################################################
#get the chromatogram files
easy_chrom = read.table('/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/20170504_QE5_nLC7_DBJ_SA_x_LFQprot_100ug_46frac_15min_15000_02.raw_Ms_BP_chromatogram.txt', header = TRUE, sep = '\t')

evo_chrom = read.table('/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/20180216_QE5_Evo1_DBJ_SA_LFQprot_HeLa_100ug_46fracs_22min_02.raw_Ms_BP_chromatogram.txt', header = TRUE, sep = '\t')

```

<br>

Plot the chromatogram data.

```{r}
###################################################################################################################################################################
#plot the chromatograms
output_plot = ggplot(easy_chrom, aes(RetentionTime, Intensity)) +
  geom_line(color = brewer.pal(6,'RdBu')[6], size = 0.75) +
  labs(x = "Retention Time (minutes)", y = 'Intensity', title = 'Easy nanoLC - Fraction 2') +
  theme(axis.text.x = element_text(size = 9), legend.position="none") +
  scale_y_continuous(limits = c(0,1.3e10), breaks = seq(0,1.3e10,2.5e9)) 
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/Ranalysis/easy-nano-chromatogram-frac2.pdf',output_plot)

output_plot = ggplot(evo_chrom, aes(RetentionTime, Intensity)) +
  geom_line(color = brewer.pal(6,'RdBu')[6], size = 0.75) +
  labs(x = "Retention Time (minutes)", y = 'Intensity', title = 'Evosep One - Fraction 2') +
  theme(axis.text.x = element_text(size = 9), legend.position="none") +
  scale_y_continuous(limits = c(0,2e9), breaks = seq(0,2e9,5e8)) 
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/Ranalysis/evo-nano-chromatogram-frac2.pdf',output_plot)


```

<br>

Now we can make a distribution of MS2 scans per cycle.


```{r}
###################################################################################################################################################################
#plot MS2 scans per cycle
easy_scan = read.table('/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/20170504_QE5_nLC7_DBJ_SA_x_LFQprot_100ug_46frac_15min_15000_02.raw_Matrix.txt', header = TRUE, sep = '\t')

output_plot = ggplot(easy_scan, aes(ParentScanRetTime, MS2ScansPerCycle)) +
  #geom_line(colour = brewer.pal(3,'PuBuGn')[3], size = 1.5, alpha = 0.75) +
  geom_point(fill = brewer.pal(6,'RdBu')[6], pch = 21, size = 3, colour = 'black', alpha = 0.2, stroke = 0.1) +
  labs(x = "Retention Time (min)", y = 'Number of Triggered MS2 Scans Per MS1', title = 'Easy nanoLC - MS2 Scans Triggering') +
  theme(axis.text.x = element_text(size = 12), legend.position="none")
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/Ranalysis/easy-nano-ms2-scans-frac2.pdf',output_plot)



evo_scan = read.table('/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/20180216_QE5_Evo1_DBJ_SA_LFQprot_HeLa_100ug_46fracs_22min_02.raw_Matrix.txt', header = TRUE, sep = '\t')

output_plot = ggplot(evo_scan, aes(ParentScanRetTime, MS2ScansPerCycle)) +
  #geom_line(colour = brewer.pal(3,'PuBuGn')[3], size = 1.5, alpha = 0.75) +
  geom_point(fill = brewer.pal(6,'RdBu')[6], pch = 21, size = 3, colour = 'black', alpha = 0.2, stroke = 0.1) +
  labs(x = "Retention Time (min)", y = 'Number of Triggered MS2 Scans Per MS1', title = 'Evosep One - MS2 Scans Triggering') +
  theme(axis.text.x = element_text(size = 12), legend.position="none")
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/Ranalysis/evo-nano-ms2-scans-frac2.pdf',output_plot)




```


<br>

This is not super informative. What if we make it as a heatmap.

```{r}
##########################################
#first we need all the easy scan data
easy_scan_files = list.files('/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/easy_scan', pattern = 'Matrix.txt', full.names = TRUE)
easy_scan_all = data.frame()
for (i in 1:length(easy_scan_files)){
  easy_temp1 = read.table(easy_scan_files[i], header = TRUE, sep = '\t')
  easy_temp2 = easy_temp1 %>% 
    dplyr::select(MS2ScanNumber, ParentScanRetTime) #select columns
  easy_temp2$fraction = i
  easy_scan_all = rbind(easy_scan_all, easy_temp2)
}

easy_scan_all$fraction = factor(easy_scan_all$fraction, levels = 1:46)
easy_scan_all$MS2Num = 1
easy_scan_all$bin_group = cut(easy_scan_all$ParentScanRetTime, breaks = seq(0,19,1))
easy_scan_compress = setDT(easy_scan_all)[,lapply(.SD,sum,na.rm=TRUE), by = c('fraction','bin_group'),.SDcols='MS2Num']
setDF(easy_scan_compress)
easy_scan_ordered = easy_scan_compress[order(easy_scan_compress$bin_group),]
easy_scan_ordered$bin_start = as.numeric(sub('\\((.*?)\\,.*','\\1',easy_scan_ordered$bin_group))
easy_scan_ordered$bin_end = as.numeric(sub('.*\\,(.*?)\\].*','\\1',easy_scan_ordered$bin_group))
easy_scan_ordered$midTime = easy_scan_ordered$bin_start + 0.5

mat1 = as.data.frame(dcast(easy_scan_ordered, fraction ~ midTime, value.var = "MS2Num"))
row.names(mat1) = mat1$fraction
mat2 = as.matrix(mat1[,c(2:20)])

#make the breaks
mat_breaks = seq(1, 1500, by = 5)
break_cols = colorRampPalette(rev(brewer.pal(n = 6, name = "RdBu")))(length(mat_breaks))


#make the actual plot
pheatmap(mat2, color = break_cols, border_color = 'white', cluster_rows = FALSE, cluster_cols = FALSE, show_colnames = TRUE, show_rownames = TRUE,
         cellwidth = 5, cellheight = 5, fontsize = 6, gaps_row = NULL, gaps_col = NULL, breaks = mat_breaks, legend = TRUE, 
         filename = '/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/Ranalysis/easy-nano-fraction-heatmap.pdf'
)

```


Now do this again for the evosep data.

```{r}
##########################################
#first we need all the easy scan data
evo_scan_files = list.files('/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/evo_scan', pattern = 'Matrix.txt', full.names = TRUE)
evo_scan_all = data.frame()
for (i in 1:length(evo_scan_files)){
  evo_temp1 = read.table(evo_scan_files[i], header = TRUE, sep = '\t')
  evo_temp2 = evo_temp1 %>% 
    dplyr::select(MS2ScanNumber, ParentScanRetTime) #select columns
  evo_temp2$fraction = i
  evo_scan_all = rbind(evo_scan_all, evo_temp2)
}

evo_scan_all$fraction = factor(evo_scan_all$fraction, levels = 1:46)
evo_scan_all$MS2Num = 1
evo_scan_all$bin_group = cut(evo_scan_all$ParentScanRetTime, breaks = seq(0,22,1))
evo_scan_compress = setDT(evo_scan_all)[,lapply(.SD,sum,na.rm=TRUE), by = c('fraction','bin_group'),.SDcols='MS2Num']
setDF(evo_scan_compress)
evo_scan_ordered = evo_scan_compress[order(evo_scan_compress$bin_group),]
evo_scan_ordered$bin_start = as.numeric(sub('\\((.*?)\\,.*','\\1',evo_scan_ordered$bin_group))
evo_scan_ordered$bin_end = as.numeric(sub('.*\\,(.*?)\\].*','\\1',evo_scan_ordered$bin_group))
evo_scan_ordered$midTime = evo_scan_ordered$bin_start + 0.5

mat1 = as.data.frame(dcast(evo_scan_ordered, fraction ~ midTime, value.var = "MS2Num"))
row.names(mat1) = mat1$fraction
mat2 = as.matrix(mat1[,c(2:22)])

#make the breaks
mat_breaks = seq(1, 1500, by = 5)
break_cols = colorRampPalette(rev(brewer.pal(n = 6, name = "RdBu")))(length(mat_breaks))


#make the actual plot
pheatmap(mat2, color = break_cols, border_color = 'white', cluster_rows = FALSE, cluster_cols = FALSE, show_colnames = TRUE, show_rownames = TRUE,
         cellwidth = 5, cellheight = 5, fontsize = 6, gaps_row = NULL, gaps_col = NULL, breaks = mat_breaks, legend = TRUE, 
         filename = '/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/Ranalysis/evo-nano-fraction-heatmap.pdf'
)

```

<br>

Do the efficiency calculations.

```{r}
###################################################################################################################################################################
#get the qc file
easy_qc = read.table('/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/easy_qc/QcDataTable.csv', header = TRUE, sep = ',')

easy_qc$acquisitionE = easy_qc$NumMs2Scans / easy_qc$TotalScans
easy_qc$injectionE = easy_qc$TotalAnalysisTime / (easy_qc$TotalAnalysisTime + 17.8)
easy_qc$MSE = easy_qc$acquisitionE * easy_qc$injectionE

mean(easy_qc$acquisitionE)
mean(easy_qc$injectionE)
mean(easy_qc$MSE)


#get the qc file
evo_qc = read.table('/projects/ptx_analysis/chughes/projects-current/morin_column-testing/published-data/evo_qc/QcDataTable.csv', header = TRUE, sep = ',')

evo_qc$acquisitionE = evo_qc$NumMs2Scans / evo_qc$TotalScans
evo_qc$injectionE = evo_qc$TotalAnalysisTime / (evo_qc$TotalAnalysisTime + 3)
evo_qc$MSE = evo_qc$acquisitionE * evo_qc$injectionE

mean(evo_qc$acquisitionE)
mean(evo_qc$injectionE)
mean(evo_qc$MSE)



```
